---
title: "NYMETS"
author: "Miranda Gutierrez"
date: "12/2/2022"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Prompt 1 Question
Prompt 1 – Using the following information provided from Baseball Savant, please provide the
following: A 100-200 word description of the pitcher’s strengths and weaknesses and a 200-300 word
recommendation to the pitching coach on what the pitcher’s development plan should be moving
forward. If there’s any information not included that would be helpful in creating that plan, note it in
your summary.

Based on the 

#Read in the dataset 
```{r}
library(ggplot2)
library(tidyverse)
library(forecast)
theme_set(theme_classic())

NY <- read.csv("~/Desktop/Mets/pitches.csv")
```

#Try to recreate the same type of graph

```{r}
ggplot(data = NY, aes(x = rel_speed, y = vert_rel_angle)) +
  coord_cartesian(xlim = c(0, 10), ylim = c(0, 100)) +
  labs(title = "MLB Players", subtitle = "NY",
       x = "REL speed", y = "spin_rate", caption = "for 2022")




ggplot(data = NY, aes(x = rel_speed, y = vert_rel_angle)) +
  
  coord_cartesian(xlim = c(0, 100), ylim = c(0, 3000)) +
  labs(title = "MLB Players", subtitle = "SF",
       x = "REL speed", y = "vert_rel_angle", caption = "for 2022")



```


```{r}

ggplot(data = NY) +
  geom_point(aes(x = rel_speed, y = vert_rel_angle))+
  geom_line(aes(x = rel_speed, y = linear) , color = "green")
  geom_line(aes(x = rel_speed, y = linear), color = "blue") +
    geom_line(aes(x = rel_speed, y = linear), color = "red")

```


```{r}
NY$pitch_type <- as.factor(NY$pitch_type)
plot(rel_speed ~ vert_rel_angle , data = NY, col = c(2,3,4,6,7)[pitch_type])
legend("topleft", col = c(2,3,4,6,7), pch = 1 ,
       legend = levels(NY$pitch_type))
      lines(NY$rel_speed,)


```

```{r}
ggplot(data = NY, aes(x=rel_speed, y=vert_rel_angle)) + geom_line(aes(colour=pitch_type))
```




```{r}
CB
CH
FF
SL

NY<- ggplot(rel_speed, aes(vert_rel_angle))
NY + geom_density(aes(fill=factor(pitch_type)), alpha=0.8) + 
    labs(title="Density plot", 
         subtitle="City Mileage Grouped by Number of cylinders",
         caption="Source: mpg",
         x="City Mileage",
         fill="# Cylinders")

```


```{r}
NY %>% ggplot(aes(x=rel_speed,))+
  geom_density()
```


```{r}
ggplot(NY, aes(x = rel_speed, y = vert_rel_angle , color = pitch_type)) +
geom_point()+
  geom_smooth()

```


```{r}



```
